Verification of a TCP/IP stack

### 2021-03-29 paper/paper.tex
Denrière partie

\ifLuaTeX
    \usepackage{fontspec}
    \IfFontExistsTF{Fira Code}
    {\setmonofont[Scale=MatchLowercase]{Fira Code}}
    {\IfFontExistsTF{DejaVu Sans Mono}{
        \setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
    }{}}
\fi

@@

@@

    The TCP connection follows an order of execution as seen in the section \ref{sec:TCP}: the connection is first opened,
    then the data are sent and received, and finally the connection can be closed. Since the sockets are an interface to use
    network protocols, they follow the same order of execution. Thus, an order can be enforced thanks to the pre- and postconditions.
    If a function $f_1$ must be executed before a function $f_2$, the strategy is to add a postcondition on the socket on $f_1$
    that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have been called first.
    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline[style=Spark]{Socket_Connect}. After its
    execution, the field \lstinline[style=Spark]{S_Remote_Ip_Addr} of the socket \lstinline[style=Spark]{Sock}. Then, since all the functions to send or receive
    data and to close the connection require a connection established, these functions have a precondition that contains a condition
    that states that the condition is established, which can be written as
    \begin{lstlisting}[style=Spark]
    Pre => Is_Ip_Initialized(Sock.S_Remote_Ip_Addr);
    \end{lstlisting}
    and because the function \lstinline{Socket_Connect} is the only one that can initialize this value, the order is respected
    if and only if the code is valid.

    The same mechanism is used with the functions \lstinline[style=Spark]{Socket_Open} that creates a valid socket when the function
    succeeds and \lstinline[style=Spark]{Socket_Connect} that requires a valid not null socket.

@@

    Most of the functions of the socket API can fail. To know if the function has failed or not, and error code is returned.
    It must be checked by the caller before continuing processing at the risk of leading to an incorrect state.
    In the previous paragraph we explained that the functions that send or receive data check in their precondition if the connection
    has been initialized or not, by checking the field \lstinline[style=Spark]{S_Remote_Ip_Addr}, and we explained that this value is initialized
    by \lstinline[style=Spark]{Socket_Connect}. The presence of the if statement \lstinline[style=Spark]{(if Error = NO_ERROR then ...)} in the postcondition
    of \lstinline[style=Spark]{Socket_Connect}  makes that \lstinline[style=Spark]{S_Remote_Ip_Addr} is initialized only when no error is reported during
    the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
    if it is known that no error has occurred previously.

    Let \lstinline[style=Spark]{Socket_Send} be the function to send data. The following example of code is incorrect because if \lstinline{Error} is
    different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}[style=Spark]
    Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
    -- GNATprove: medium: precondition might fail.
    Socket_Send(Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    In comparison this second example of code does not produce any check message because when \lstinline[style=Spark]{Socket_Send} is called, GNATprove knows that
    \lstinline[style=Spark]{ERROR} is equal to \lstinline[style=Spark]{NO_ERROR} and then \lstinline[style=Spark]{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}[style=Spark]
    Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
    if Error = NO_ERROR then
        Socket_Send(Sock, Data, Written, Flags, Error);
    end if;
    \end{lstlisting}

    The same mechanism can be observed on figure \ref{Fig:socketInterface:SPARK}. The user has to test if the socket returned by
    \lstinline[style=Spark]{Socket_Open} is null or not before calling \lstinline[style=Spark]{Socket_Connect}, otherwise GNATprove will emit a message.

@@

    The parameter \textit{Event\_Mask} represents the event that we are waiting for. A loop of three steps simulates the
    reception of three messages with a call to the function \lstinline[style=Spark]{Tcp_Process_One_Segment}. The new events
    satisfied are computed, and if the searched event is reached, the socket is returned.
    Then, it is possible for SPARK to compute all the relations that link the state of the input socket with the states
    of the returned socket, depending on the event waited.
    The loop is limited to three steps because all the paths in the automaton that do not require a user action have
    a length of at most three.

@@

    The work done on the TCP implementation has shown the powerful
    of SPARK in finding bugs. One of the bugs found, a memory leak,
    relies on the use of the pointers.

>>>>

@@

\let\Spark\lstinline

@@

    The TCP connection follows an order of execution as seen in the section \ref{sec:TCP}: the connection is first opened,
    then the data are sent and received, and finally the connection can be closed. Since the sockets are an interface to use
    network protocols, they follow the same order of execution. Thus, an order can be enforced thanks to the pre- and postconditions.
    If a function $f_1$ must be executed before a function $f_2$, the strategy is to add a postcondition on the socket on $f_1$
    that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have been called first.
    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \Spark{Socket_Connect}. After its
    execution, the field \Spark{S_Remote_Ip_Addr} of the socket \Spark{Sock}. Then, since all the functions to send or receive
    data and to close the connection require a connection established, these functions have a precondition that contains a condition
    that states that the condition is established, which can be written as
    \begin{lstlisting}[style=Spark]
    Pre => Is_Ip_Initialized(Sock.S_Remote_Ip_Addr);
    \end{lstlisting}
    and because the function \Spark{Socket_Connect} is the only one that can initialize this value, the order is respected
    if and only if the code is valid.

    The same mechanism is used with the functions \Spark{Socket_Open} that creates a valid socket when the function
    succeeds and \Spark{Socket_Connect} that requires a valid not null socket.

@@

    Most of the functions of the socket API can fail. To know if the function has failed or not, and error code is returned.
    It must be checked by the caller before continuing processing at the risk of leading to an incorrect state.
    In the previous paragraph we explained that the functions that send or receive data check in their precondition if the connection
    has been initialized or not, by checking the field \Spark{S_Remote_Ip_Addr}, and we explained that this value is initialized
    by \Spark{Socket_Connect}. The presence of the if statement \Spark{(if Error = NO_ERROR then ...)} in the postcondition
    of \Spark{Socket_Connect}  makes that \Spark{S_Remote_Ip_Addr} is initialized only when no error is reported during
    the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
    if it is known that no error has occurred previously.

    Here, \Spark{Socket_Send} refers to the function to send data. The following example of code is incorrect because if \lstinline{Error} is
    different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}[style=Spark]
    Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
    -- GNATprove: medium: precondition might fail.
    Socket_Send(Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    In comparison this second example of code does not produce any check message because when \Spark{Socket_Send} is called, GNATprove knows that
    \Spark{ERROR} is equal to \Spark{NO_ERROR} and then \Spark{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}[style=Spark]
    Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
    if Error = NO_ERROR then
        Socket_Send(Sock, Data, Written, Flags, Error);
    end if;
    \end{lstlisting}

    The same mechanism can be observed on figure \ref{Fig:socketInterface:SPARK}. The user has to test if the socket returned by
    \Spark{Socket_Open} is null or not before calling \Spark{Socket_Connect}, otherwise GNATprove will emit a message.

@@

    The parameter \textit{Event\_Mask} represents the event that we are waiting for. A loop of three steps simulates the
    reception of three messages with a call to the function \Spark{Tcp_Process_One_Segment}. The new events
    satisfied are computed, and if the searched event is reached, the socket is returned.
    Then, it is possible for SPARK to compute all the relations that link the state of the input socket with the states
    of the returned socket, depending on the event waited.
    The loop is limited to three steps because all the paths in the automaton that do not require a user action have
    a length of at most three.

@@

    The work done on the TCP implementation has shown the strength
    of SPARK in finding bugs. One of the bugs found, a memory leak,
    relies on the use of the pointers.

\subsubsection{Concurrency bug}

    The most significant problem found in the TCP implementation relies on the concurrency.
    Concurrency is one of the biggest source of bug when writing code. It is hard to understand
    all the possible interactions between different threads which leads to many bugs that can be
    difficult to reproduce and to debug. The use of SPARK helps us to find one.

    %% TODO : Show the relevant code snippet here

\subsubsection{Memory leak}

    A memory leak has been found thanks by SPARK and the recent work of Dross and Kanig~\cite{dross2020recursive}.
    They added a pointer support to SPARK inspired by Rust. At any program point, every memory cell must be owned by a pointer.
    If a memory cell has not be freed and is not owned anymore by any pointer, there is a memory leaks.
    This mechanism of ownership has helped in finding one memory leak in the original C code that has been fixed in the
    SPARK version of the code.

%% Ici on peut parler d'une ou deux typos trouvées dans le code. Mais plutôt liées à une relecture de code et pas vraiment dues à SPARK

\subsubsection{Metrics of the new implementation}

\subsubsection{Tests to validate the new implementation}
